CFF GCT Study

### 2025-05-29 Christine Chan/CFF GCT Study/cff_gct_analysis.qmd
more plots for Larry

```{r setup}
#| include: false
library(readxl)
library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)
library(ggpmisc)
library(pROC)
library(rspiro)
library(childsds)
library(arsenal)
library(gt)
library(reporttools)
library(rpart)
library(rpart.plot)
library(rms)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
  "Linux" = "/home/timvigers/OneDrive/Vigers/CF"
)
github_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Documents/GitHub",
  "Linux" = "/home/timvigers/Documents/GitHub"
)
knitr::opts_knit$set(root.dir = home_dir)
```

@@

```{r}
#| label: fig-boxplot
#| fig-cap: Boxplot of OGTT glucose values by diabetes status
#| warning: false
plot_df <- df %>%
  pivot_longer(c(`OGTT 0 min`, `OGTT 60 min`, `OGTT 120 min`),
    names_to = "OGTT Timepoint", values_to = "Glucose"
  )
plot_df$`OGTT Timepoint` <- factor(plot_df$`OGTT Timepoint`,
  levels = c("OGTT 0 min", "OGTT 60 min", "OGTT 120 min")
)
box_p <- ggplot(
  plot_df,
  aes(x = `OGTT Timepoint`, y = Glucose, fill = `OGTT Timepoint`)
) +
  geom_boxplot() +
  scale_y_continuous(breaks = c(100, 140, 200, 300)) +
  facet_wrap(~Diabetes) +
  theme_bw(base_size = 20) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title.x = element_blank()
  )
box_p
ggsave("./Christine Chan/CFF GCT Study/Dissemination/Figures/boxplot.png",
  box_p,
  device = "png"
)
# OGTT 1 hour and GCT 1 hour regression
scatter_p <- ggplot(df, aes(x = `GCT 60 min plasma`, y = `OGTT 60 min`)) +
  geom_point() +
  stat_poly_line() +
  stat_poly_eq(use_label(c("eq", "R2")),
    size = 5,
    label.y = "bottom", label.x = "right"
  ) +
  theme_bw(base_size = 20)
scatter_p
ggsave("./Christine Chan/CFF GCT Study/Dissemination/Figures/scatterplot.png",
  scatter_p,
  device = "png"
)
```

@@

```{r}
f <- lrm(CFRD ~ `GCT 0 min capillary` + HbA1c,
  data = df
)
ddist <- datadist(df)
options(datadist = "ddist")
nom <- nomogram(f, fun = plogis, funlabel = "Risk of CFRD",lp = F)
plot(nom)
```

@@

```{r}
f <- lrm(CFRD ~ `GCT 0 min capillary` + HbA1c + `GCT 60 min plasma`,
  data = df
)
ddist <- datadist(df)
options(datadist = "ddist")
nom <- nomogram(f, fun = plogis, funlabel = "Risk of CFRD",lp = F)
plot(nom)
```

@@

```{r}
f <- lrm(CFRD ~ `GCT 0 min capillary` + `GCT 60 min plasma`,
  data = df
)
ddist <- datadist(df)
options(datadist = "ddist")
nom <- nomogram(f, fun = plogis, funlabel = "Risk of CFRD",lp = F)
plot(nom)
```

>>>>

```{r setup}
#| include: false
library(readxl)
library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)
library(ggpmisc)
library(pROC)
library(rspiro)
library(childsds)
library(arsenal)
library(gt)
library(reporttools)
library(rpart)
library(rpart.plot)
library(rms)
library(patchwork)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
  "Linux" = "/home/tim/OneDrive/Vigers/CF"
)
github_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/tim/Documents/GitHub",
  "Windows" = "C:/Users/Tim/Documents/GitHub",
  "Linux" = "/home/tim/Documents/GitHub"
)
knitr::opts_knit$set(root.dir = home_dir)
```

@@

```{r}
#| label: fig-boxplot
#| fig-cap: Boxplot of OGTT glucose values by diabetes status
# The code below is a total mess because Larry keeps sending requests -
# clean up later
plot_df <- df %>%
  pivot_longer(c(`OGTT 0 min`, `OGTT 60 min`, `OGTT 120 min`),
    names_to = "OGTT Timepoint", values_to = "Glucose"
plot_df$`OGTT Timepoint` <- factor(plot_df$`OGTT Timepoint`,
  levels = c("OGTT 0 min", "OGTT 60 min", "OGTT 120 min")
box_p <- ggplot(
  plot_df,
  aes(x = `OGTT Timepoint`, y = Glucose, fill = `OGTT Timepoint`)
) +
  geom_boxplot() +
  scale_y_continuous(breaks = c(100, 140, 200, 300)) +
  facet_wrap(~Diabetes) +
  theme_bw(base_size = 20) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title.x = element_blank()
box_p
ggsave("./Christine Chan/CFF GCT Study/Dissemination/Figures/boxplot.png",
  box_p,
  device = "png"
# Dotplot
dotplot <- ggplot(df, aes(x = Diagnosis, y = `GCT 60 min plasma`)) +
  geom_point() +
  theme_bw()
dotplot
dotplot +
  theme(text = element_blank())
# Additional boxplots/dotplot per Christine and Larry
plot_df <- df %>%
  pivot_longer(c(`OGTT 0 min`, `OGTT 60 min`, `OGTT 120 min`, `GCT 60 min plasma`),
    names_to = "OGTT Timepoint", values_to = "Glucose"
  filter(`OGTT Timepoint` != "OGTT 60 min")
plot_df$`OGTT Timepoint` <- factor(plot_df$`OGTT Timepoint`,
  levels = c("OGTT 0 min", "GCT 60 min plasma", "OGTT 120 min")
box_p <- ggplot(
  plot_df,
  aes(x = `OGTT Timepoint`, y = Glucose, fill = `OGTT Timepoint`)
) +
  geom_boxplot() +
  scale_y_continuous(breaks = c(100, 140, 200, 300)) +
  facet_wrap(~Diabetes) +
  theme_bw(base_size = 20) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title.x = element_blank()
box_p
ggsave("./Christine Chan/CFF GCT Study/Dissemination/Figures/boxplot.png",
  box_p,
  device = "png"
# GCTpl, RCG, and HbA1c
plot_df <- df %>%
  pivot_longer(c(`GCT 60 min plasma`, `GCT 0 min capillary`, HbA1c),
    names_to = "Measure", values_to = "Value"
box_p1 <- ggplot(
  plot_df[plot_df$Measure == "GCT 60 min plasma", ],
  aes(x = Diabetes, y = Value, fill = Diabetes)
) +
  geom_boxplot() +
  geom_hline(yintercept = 146.5, linetype = "dashed") +
  ylim(c(50, 300)) +
  ylab("mg/dL") +
  ggtitle("GCTpl") +
  theme_bw(base_size = 15) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title.x = element_blank()
box_p2 <- ggplot(
  plot_df[plot_df$Measure == "GCT 0 min capillary", ],
  aes(x = Diabetes, y = Value, fill = Diabetes)
) +
  geom_boxplot() +
  geom_hline(yintercept = 98.5, linetype = "dashed") +
  ylim(c(50, 300)) +
  ggtitle("RCG") +
  theme_bw(base_size = 15) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title.x = element_blank(),
    axis.title.y = element_blank()
box_p3 <- ggplot(
  plot_df[plot_df$Measure == "HbA1c", ],
  aes(x = Diabetes, y = Value, fill = Diabetes)
) +
  geom_boxplot() +
  geom_hline(yintercept = 5.45, linetype = "dashed") +
  ylab("%") +
  ggtitle("HbA1c") +
  theme_bw(base_size = 15) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title.x = element_blank(),
    axis.title.y = element_blank()
box_p <- box_p1 + box_p2 + box_p3
ggsave("./Christine Chan/CFF GCT Study/Dissemination/Figures/boxplot2.png",
  box_p,
  device = "png"
# Dotplot
plot_df <- df %>%
  pivot_longer(c(`GCT 60 min plasma`, `GCT 0 min capillary`, HbA1c),
    names_to = "Measure", values_to = "Value"
box_p1 <- ggplot(
  plot_df[plot_df$Measure == "GCT 60 min plasma", ],
  aes(x = Diabetes, y = Value, fill = Diabetes, colour = Diabetes)
) +
  geom_jitter() +
  geom_hline(yintercept = 146.5, linetype = "dashed") +
  ylim(c(50, 300)) +
  ylab("mg/dL") +
  ggtitle("GCTpl") +
  theme_bw(base_size = 15) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title.x = element_blank()
box_p2 <- ggplot(
  plot_df[plot_df$Measure == "GCT 0 min capillary", ],
  aes(x = Diabetes, y = Value, fill = Diabetes, colour = Diabetes)
) +
  geom_jitter() +
  geom_hline(yintercept = 98.5, linetype = "dashed") +
  ylim(c(50, 300)) +
  ggtitle("RCG") +
  theme_bw(base_size = 15) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title.x = element_blank(),
    axis.title.y = element_blank()
box_p3 <- ggplot(
  plot_df[plot_df$Measure == "HbA1c", ],
  aes(x = Diabetes, y = Value, fill = Diabetes, colour = Diabetes)
) +
  geom_jitter() +
  geom_hline(yintercept = 5.45, linetype = "dashed") +
  ylab("%") +
  ggtitle("HbA1c") +
  theme_bw(base_size = 15) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title.x = element_blank(),
    axis.title.y = element_blank()
box_p <- box_p1 + box_p2 + box_p3
ggsave("./Christine Chan/CFF GCT Study/Dissemination/Figures/dotplot2.png",
  box_p,
  device = "png"
# OGTT 1 hour and GCT 1 hour regression
scatter_p <- ggplot(df, aes(x = `GCT 60 min plasma`, y = `OGTT 60 min`)) +
  geom_point() +
  stat_poly_line() +
  stat_poly_eq(use_label(c("eq", "R2")),
    size = 5,
    label.y = "bottom", label.x = "right"
  ) +
  theme_bw(base_size = 20)
scatter_p
ggsave("./Christine Chan/CFF GCT Study/Dissemination/Figures/scatterplot.png",
  scatter_p,
  device = "png"

@@

```{r}
f <- lrm(CFRD ~ `GCT 0 min capillary` + HbA1c,
  data = df
)
ddist <- datadist(df)
options(datadist = "ddist")
nom <- nomogram(f, fun = plogis, funlabel = "Risk of CFRD", lp = F)
plot(nom)
```

@@

```{r}
f <- lrm(CFRD ~ `GCT 0 min capillary` + HbA1c + `GCT 60 min plasma`,
  data = df
)
ddist <- datadist(df)
options(datadist = "ddist")
nom <- nomogram(f, fun = plogis, funlabel = "Risk of CFRD", lp = F)
plot(nom)
```

@@

```{r}
f <- lrm(CFRD ~ `GCT 0 min capillary` + `GCT 60 min plasma`,
  data = df
)
ddist <- datadist(df)
options(datadist = "ddist")
nom <- nomogram(f, fun = plogis, funlabel = "Risk of CFRD", lp = F)
plot(nom)
```
